\section{Introduction}

Data centers (\DCs) are a critical piece of today's computing infrastructure
that support key Internet applications.  In this context, \DC  network designs
must satisfy several potentially conflicting goals---performance (e.g.,
minimize oversubscribed links, low latency)~\cite{fattree,vl2}, equipment and
management cost~\cite{fattree,popa-cost}, flexibility to adapt to changing
traffic patterns~\cite{proteus,3db,flyways,cthru}, incremental expandability to
add new servers or racks~\cite{legup,jellyfish}, and other practical concerns
including cabling complexity, power, and
cooling~\cite{farrington,portland,cabling}.

Given the highly variable and unpredictable nature of \DC
workloads~\cite{vl2}, early \DC designs offered extreme points
in the space of cost-performance tradeoffs---either poor performance
at low cost (e.g., a simple tree has many oversubscribed links) or
expensive over-provisioned solutions with good worst-case performance
(e.g., fat-trees for full bisection bandwidth~\cite{fattree}). Recent
works suggest a middle ground that dynamically augments a simple
fixed infrastructure with additional inter-rack
wireless~\cite{flyways,3db} or optical links~\cite{cthru} to alleviate
congested hotspots.  While these do offer some 
 performance benefits, we believe that they do not push the envelop far
enough; e.g., they continue to incur high cabling complexity 
 and may only handle specific traffic patterns.

\eat{While these do offer 
 performance benefits, they also introduce additional 
challenges. RF wireless links introduce significant interference 
and hence have limited performance benefit.  Optical links 
suffer from scalability, limited flexibility, and full-switch 
latency.}

% early datacenter designers were initially 


%\footnote{While
%  optical interconnects have also been suggested~\cite{cthru}, the use
%  of wireless is more appealing as it does not burden operators with
%  cabling costs.} %\red{I would just move the footnote inline ..} 

%% story flow is broken now with logical extreme etc.. just making it more to the pt
In this work, we explore the vision 
of a \emph{all-wireless inter-rack \DC fabric}.  This vision,
if realized, would provide unprecedented degrees of flexibility for
\DCs.  For example, it will allow operators to dynamically
reconfigure the {\em entire} network topology to adapt to changing
traffic demands.  Similarly, it can act as an enabler for operators to
  deploy topologies  that would otherwise remain
``paper designs'' due to the perceived cabling complexity.

%(e.g., random graphs that offer better incremental expandability~\cite{jellyfish}),
%\vyas{something else}

Unfortunately, existing wireless/RF technologies are not suitable on
two fronts. First, they incur a large interference footprint even with
advanced phased-array antennas, especially when side lobes are
considered~\cite{cornell}. Second, they suffer from a significant
drop-off in data rates at longer distances~\cite{3db}, as 
 federal regulations prevent use of higher power or wider
bandwidth.  In conjunction, these factors fundamentally limit the
number and types of inter-rack links that can be created---an issue
that exists even with newer designs to extend the range via ceiling
reflectors~\cite{3db}.

Consequently, we look beyond traditional RF-based solutions and
explore a somewhat non-standard wireless technology, namely {\em
  Free-Space Optics} (FSO).  FSO uses visible or infra-red lasers to
implement point-to-point data links, at very high data rates
($>$1~Gbps) and at longer ranges. (We elaborate on these
in \Section\ref{sec:fso}.) While the use of FSO in a general
communications context is not new, there has been very little work to
systematically explore the viability of FSO in the \DC or
highlight the benefits that FSO offers in this context.\footnote{The
  only parallel work we are aware of is a recent patent
  document~\cite{us-patent}. Unfortunately, this offers little in
  terms of viability analysis, design space arguments, or
  performance tradeoffs.}


While FSO is an enabler, there are two significant challenges that need to be
addressed.  First, off-the-shelf FSO  links cannot be
cost-effectively reconfigured (i.e., realigned) at fast time-scales necessary
in the DC context. Second, cost and physical constraints limit the number of
FSO links that can be installed on the top of each rack.  Thus, we need an
effective topology design and management framework that can provide the desired
performance and flexibility while operating under the physical and cost
constraints.


In this paper, we discuss the viability of an FSO-based all-wireless inter-rack
fabric and present early solutions to address the above challenges.  To achieve
fast reconfigurability, we leverage {\em switchable mirrors} that can be
electronically controlled to act as mirrors or pass-through glasses.  
 The use of switchable mirrors 
 avoids the need for careful realignment (\S\ref{sec:design}). We also
discuss topology design and reconfiguration heuristics that seem to perform
well in practice.  We have built a proof-of-concept prototype using
off-the-shelf FSO and switchable mirror solutions. Our early  simulation
 results show that our approach provides superior performance  compared to 
state-of-art \DC networks of comparable cost (\Section\ref{sec:perf}).
We discuss extensions to our basic approach in \S\ref{sec:future}, before
concluding in \S\ref{sec:conc}.  We discuss related work inline throughout the
paper.











